---
title: "combine combustion data by project (FISL, LC, LCC"
author: "Anna Talucci"
date: "2025-06-06"
output: html_document
---
```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

Combine CSV files at project level for combustion model and for day of burn data.

# Packages 

```{r}
library(tidyverse)
library(sf)
```


# Data 

# Points

```{r}
lc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/lc_pts_WGS84_2025-06-02.shp", "lc_pts_WGS84_2025-06-02")
```

```{r}
lcc_pts = st_read("../outputs/LCC_cleaned/points/shapefiles/lcc_pts_WGS84_2025-06-02.shp", "lcc_pts_WGS84_2025-06-02")
```

```{r}
fisl_pts =st_read("../outputs/LCC_cleaned/points/shapefiles/fisl_pts_WGS84_2025-06-02.shp", "fisl_pts_WGS84_2025-06-02")
```

# CSV

## FISL
```{r}
fisl_oldrdnbr_list = list.files("../data/combustion_data_layers/FISL/old_rdnbr", pattern='.csv$', full.names = TRUE)
fisl_oldrdnbr_list
```

```{r}
fisl_indices_list = list.files("../data/combustion_data_layers/FISL/indices", pattern='.csv$', full.names = TRUE)
fisl_indices_list
```

```{r}
fisl_tct_list = list.files("../data/combustion_data_layers/FISL/tct", pattern='.csv$', full.names = TRUE)
fisl_tct_list
```

```{r}
fisl_treecover_list = list.files("../data/combustion_data_layers/FISL/treecover", pattern='.csv$', full.names = TRUE)
fisl_treecover_list
```

```{r}
fisl_static_list = list.files("../data/combustion_data_layers/FISL/static", pattern='.csv$', full.names = TRUE)
fisl_static_list
```

```{r}
fisl_dob_list = list.files("../data/combustion_data_layers/FISL/dob", pattern='.csv$', full.names = TRUE)
fisl_dob_list
```

```{r}
fisl_cna = read_csv("../data/combustion_data_layers/FISL/2025-06-02_climateNA_FISL_Normal_1981_2010MSY.csv")
```
```{r}
fisl_fwi_gfwed = read_csv("../outputs/LCC_cleaned/csv/2025-06-23_Fisl_fwi_gfwed.csv")
```

## LC

```{r}
lc_oldrdnbr_list = list.files("../data/combustion_data_layers/LC/old_rdnbr", pattern='.csv$', full.names = TRUE)
lc_oldrdnbr_list
```

```{r}
lc_indices_list = list.files("../data/combustion_data_layers/LC/indices", pattern='.csv$', full.names = TRUE)
lc_indices_list
```

```{r}
lc_tct_list = list.files("../data/combustion_data_layers/LC/tct", pattern='.csv$', full.names = TRUE)
lc_tct_list
```

```{r}
lc_treecover_list = list.files("../data/combustion_data_layers/LC/treecover", pattern='.csv$', full.names = TRUE)
lc_treecover_list
```

```{r}
lc_static_list = list.files("../data/combustion_data_layers/LC/static", pattern='.csv$', full.names = TRUE)
lc_static_list
```

```{r}
lc_dob_list = list.files("../data/combustion_data_layers/LC/dob", pattern='.csv$', full.names = TRUE)
lc_dob_list
```

```{r}
lc_cna = read_csv("../data/combustion_data_layers/LC/2025-06-02_climateNA_LC_Normal_1981_2010MSY.csv")
```
```{r}
lc_fwi_gfwed = read_csv("../outputs/LCC_cleaned/csv/2025-06-23_LC_fwi_gfwed.csv")
```


## ORG

```{r}
org_indices_list = list.files("../data/combustion_data_layers/ORG/indices", pattern='.csv$', full.names = TRUE)
org_indices_list
```

## Field data

```{r}
lc_field = read_csv("../outputs/LCC_cleaned/csv/LC_ID_2025-06-02.csv")

fisl_field = read_csv("../outputs/LCC_cleaned/csv/FISL_ID_2025-06-02.csv")
```

# Read list of files 

## Functions for csv list

```{r}
csv_rbind <- function(file_list) {
  csv_data <- purrr::map_dfr(file_list, readr::read_csv)
  return(csv_data)
}

```

```{r}
csv_cbind <- function(file_list, id_col = "ID") {
  
  # Read and join using reduce with left_join by ID
  data_list <- purrr::map(file_list, ~readr::read_csv(.x, show_col_types = FALSE))
  
  # Ensure ID column exists in all files
  data_list <- purrr::map(data_list, ~dplyr::select(., dplyr::all_of(id_col), dplyr::everything()))
  
  # Join all data frames by ID
  joined_data <- purrr::reduce(data_list, ~dplyr::full_join(.x, .y, by = id_col))
  
  return(joined_data)
}

```




## dob

```{r}
( fisl_dob = csv_rbind(fisl_dob_list) ) 
( lc_dob = csv_rbind(lc_dob_list) ) 

```


## indices
```{r}
( org_indices = csv_rbind(org_indices_list) %>% select(id, rdnbr)) 
```

```{r}
( fisl_indices = csv_rbind(fisl_indices_list) ) 
( lc_indices = csv_rbind(lc_indices_list) ) 

```


## tree cover

```{r}
( fisl_treecover = csv_rbind(fisl_treecover_list) ) 
( lc_treecover = csv_rbind(lc_treecover_list) ) 

```


## tct

```{r}
( fisl_tct = csv_rbind(fisl_tct_list) ) 
( lc_tct = csv_rbind(lc_tct_list) ) 
```

# Static

```{r}
( fisl_static = csv_cbind(fisl_static_list))
( lc_static = csv_cbind(lc_static_list))
```



# clean data

## functions

```{r}
clean_rbind = function(df){
  df %>%
    select(-.geo) %>%
    select(-`system:index`) %>%
    relocate(ID)
}
```


```{r}
clean_cbind = function(df){
  df %>%
    select(-contains(".y")) %>%
    select(-contains(".x")) 
}
```

## dob

```{r}
( fisl_dob_clean = clean_rbind(fisl_dob) )
( lc_dob_clean = clean_rbind(lc_dob) )

```



## indices

```{r}
( fisl_indices_clean = clean_rbind(fisl_indices) )
( lc_indices_clean = clean_rbind(lc_indices) )

```
## tct

```{r}
( fisl_tct_clean = clean_rbind(fisl_tct) )
( lc_tct_clean = clean_rbind(lc_tct) )

```

## treecover

```{r}
( fisl_treecover_clean = clean_rbind(fisl_treecover) )
( lc_treecover_clean = clean_rbind(lc_treecover) )

```

## static

```{r}
( fisl_static_clean = clean_cbind(fisl_static) )
( lc_static_clean = clean_cbind(lc_static) )

```

## CNA


```{r}
clean_cna = function(df){
  df %>% 
  rename(ID = id1) %>%
    select(ID, AHM, bFFP, CMD, CMD05, CMD06, CMD07, CMD08, DD_0, DD_0_05, DD_0_06, DD_0_07, DD_0_08, DD_18, DD18_05, DD18_06, DD18_07, DD18_08, DD18, DD18_sm, DD18_05, DD18_06, DD18_07, DD18_08, DD5, DD5_05, DD5_06, DD5_07, DD5_08, eFFP, EMT, Eref, Eref05, Eref06, Eref07, Eref08, EXT, FFP, MAP, MAR, MAT, MCMT, MSP, MWMT, NFFD, NFFD05, NFFD06, NFFD07, NFFD08, PAS, PAS05, PAS06, PAS07, PAS08, PPT05, PPT06, PPT07, PPT08, Rad05, Rad06, Rad07, Rad08, RH, RH05, RH06, RH07, RH08, SHM, Tave05, Tave06, Tave07, Tave08, TD, Tmax05, Tmax06, Tmax07, Tmax08, Tmin05, Tmin06, Tmin07, Tmin08) %>%
    mutate(CMD_5_8 = ((CMD05 + CMD06 + CMD07 + CMD08)/4),
           DD_0_5_8 = ((DD_0_05 + DD_0_06 + DD_0_07 + DD_0_08)/4),
           DD_18_5_8 = ((DD18_05 + DD18_06 + DD18_07 + DD18_08)/4),
           DD5_5_8 = ((DD5_05 + DD5_06 + DD5_07 + DD5_08)/4),
           Eref_5_8 = ((Eref05 + Eref06 + Eref07 + Eref08)/4),
           NFFD_5_8 = ((NFFD05 + NFFD06 + NFFD07 + NFFD08)/4),
           PAS_5_8 = ((PAS05 + PAS06 + PAS07 + PAS08)/4),
           PPT_5_8 = ((PPT05 + PPT06 + PPT07 + PPT08)/4),
           Rad_5_8 = ((Rad05 + Rad06 + Rad07 + Rad08)/4), 
           RH_5_8 = ((RH05 + RH06 + RH07 + RH08)/4),
           Tave_5_8 = ((Tave05 + Tave06 + Tave07 + Tave08)/4),
           Tmax_5_8 = ((Tmax05 + Tmax06 + Tmax07 + Tmax08)/4),
           Tmin_5_8 = ((Tmin05 + Tmin06 + Tmin07 + Tmin08)/4)) %>%
    select(-contains("05"), -contains("06"), -contains("07"), -contains("08")) %>%
     dplyr::rename_with(~ paste("CNA", .x, sep = "_")) %>%
    rename(ID = CNA_ID)
}
```


```{r}
(fisl_cna_clean = clean_cna(fisl_cna) )
( lc_cna_clean = clean_cna(lc_cna) )

```
```{r}
sort(names(fisl_cna_clean))
```
("AHM", "bFFP", "CMD", "CMD05", "CMD06", "CMD07", "CMD08", "DD_0", "DD_0_05", "DD_0_06", "DD_0_07", "DD_0_08", "DD_18", "DD18_05", "DD18_06", "DD18_07", "DD18_08", "DD18", "DD18_05",   "DD18_06", "DD18_07", "DD18_08", "DD5",  "DD5_05", "DD5_06", "DD5_07", "DD5_08", "eFFP", "EMT", "Eref", "Eref05", "Eref06", "Eref07", "Eref08", "EXT", "FFP", "MAP", "MAR", "MAT","MCMT", "MSP", "MWMT", "NFFD", "NFFD05", "NFFD06",    "NFFD07", "NFFD08", "PAS", "PAS05", "PAS06", "PAS07", "PAS08", "PPT05", "PPT06", "PPT07", "PPT08", "Rad05", "Rad06", "Rad07", "Rad08", "RH", "RH05", "RH06", "RH07", "RH08",  "SHM", "Tave05", "Tave06", "Tave07", "Tave08", "TD", "Tmax05", "Tmax06", "Tmax07", "Tmax08", "Tmin05", "Tmin06", "Tmin07", "Tmin08" )
AHM
bFFP
CMD
CMD_5_8 (need individual months)
DD_0
DD_0_5_8
DD_18
DD_18_5_8
DD18
DD18_sm
DD5
DD5_5_8
eFFP
EMT
Eref
Eref_5_8
EXT
FFP
MAP
MAR
MAT
MCMT
MSP
MWMT
NFFD
NFFD_5_8
PAS
PAS_5_8
PPT_5_8
Rad_5_8
RH
RH_5_8
SHM
Tave_5_8
TD
Tmax_5_8
Tmin_5_8

# Original point files as df
```{r}
lc_field
fisl_field
```

```{r}
( fisl_combusted = fisl_field %>% 
  select(ID, comb.below, comb.trees, comb.depth) %>%
  rename(above.carbon.combusted = comb.trees, below.ground.carbon.combusted = comb.below, burn.depth = comb.depth) )
```

```{r}
( lc_combusted = lc_field %>% 
  select(ID, combustion_above, burn_depth, total_c_combusted) %>%
  mutate(below.ground.carbon.combusted = total_c_combusted-combustion_above) %>%
  rename(above.carbon.combusted = combustion_above, burn.depth = burn_depth) )
```
above.carbon.combusted
below.ground.carbon.combusted
```{r}
( fisl_pts_df = fisl_pts %>% st_drop_geometry() )
( lc_pts_df = lc_pts %>% st_drop_geometry() )

```

# Combine orginal, static, indices, tct, and tree cover by project

## Function

```{r}
combine_project_data  = function(pts, static, indices, tct, treecover, cna, fwi_gfwed) {
  join_data = full_join(pts, static, by="ID") %>%
    full_join(., indices, by="ID") %>%
    full_join(., tct, by="ID") %>%
    full_join(., treecover, by="ID") %>%
    full_join(., cna, by="ID") %>%
    full_join(., fwi_gfwed, by="ID") %>%
    select(-fireYr) %>%
    rename(lat=lat.x, lon=lon.x, fireYr = fireYr.x, NDII = dNDII, NDVI = dNDVI, Relative.humidity = rh, Temperature =t, Tree.cover = tree_canopy_cover, VPD = vpd, Wind.speed = wspeed) %>%
    relocate(ID, lat, lon, fireYr) %>%
    select(-contains(".y")) %>%
    select(-contains(".x"))
}
```

```{r}
( fisl_df = combine_project_data(fisl_pts_df, fisl_static_clean, fisl_indices_clean, fisl_tct_clean, fisl_treecover_clean, fisl_cna_clean, fisl_fwi_gfwed) )

( lc_df = combine_project_data(lc_pts_df, lc_static_clean, lc_indices_clean, lc_tct_clean, lc_treecover_clean, lc_cna_clean, lc_fwi_gfwed) )


```

## Add field data

```{r}
(fisl_df1 = full_join(fisl_df, fisl_combusted, by = "ID") )
(lc_df1 = full_join(lc_df, lc_combusted, by = "ID") )
```

## Variable names

```{r}
sort(names(fisl_df1))
```

```{r}
sort(names(lc_df1))
```

above.carbon.combusted
below.ground.carbon.combusted
burn_year
Date
DSR

## Save LC and FISL for combustion model


```{r}
write_csv(lc_df1, "../outputs/LCC_cleaned/csv/2025-07-02_LC_forModel.csv")
write_csv(fisl_df1, "../outputs/LCC_cleaned/csv/2025-07-02_FISL_forModel.csv")
```

```{r}
write_csv(org_indices, "../outputs/LCC_cleaned/csv/2025-07-03_ORG_rdnbr.csv")
```



## missing data for combustion variable data frames

```{r}
( fisl_with_na <- fisl_df %>% filter(if_any(everything(), is.na)) )

( lc_with_na <- lc_df %>% filter(if_any(everything(), is.na)) )
```

# Missing data for DOB

## Project

### Disnct ids

```{r}
n_distinct(fisl_df$ID)
n_distinct(lc_df$ID)

```

### show rows with na

```{r}
( fisl_with_na <- fisl_df %>% filter(if_any(everything(), is.na)) )

( lc_with_na <- lc_df %>% filter(if_any(everything(), is.na)) )

( lcc_with_na <- lcc_df %>% filter(if_any(everything(), is.na)) )
```
### drop na



## Dob
## Function

```{r}
combine_org_dob_data  = function(pts, dob) {
  join_data = full_join(pts, dob, by="ID") 
}
```

```{r}
( fisl_dob_df = combine_org_dob_data(fisl_pts, fisl_dob_clean) )

( lc_dob_df = combine_org_dob_data(lc_pts, lc_dob_clean) )

( lcc_dob_df = combine_org_dob_data(lcc_pts, lcc_dob_clean) )
```


```{r}
( fisl_dob_with_na <- fisl_dob_df %>% 
    filter(if_any(everything(), is.na)) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lc_dob_with_na <- lc_dob_df %>% 
    filter(if_any(everything(), is.na))%>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lcc_dob_with_na <- lcc_dob_df %>% 
    filter(if_any(everything(), is.na)) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")) %>%
    mutate(fireYr = as.numeric(fireYr)))
```

```{r}
fisl_dob_with_na
lc_dob_with_na
```
```{r}
lcc_dob_with_na
```
```{r}
sort(unique(fisl_dob_with_na$fireYr))
```

## Save to geopackage

```{r eval=FALSE, include=FALSE}
st_write(fisl_dob_with_na, "../outputs/dob_na/fisl_dob_na_2025-06-06.gpkg")
st_write(lc_dob_with_na, "../outputs/dob_na/lc_dob_na_2025-06-06.gpkg")
st_write(lcc_dob_with_na, "../outputs/dob_na/lcc_dob_na_2025-06-06.gpkg")
```

# DOB for FWI

Drop NA and save



```{r}
( fisl_dob_drop_na <- fisl_dob_df %>% 
    drop_na(BurnDate) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lc_dob_drop_na <- lc_dob_df %>% 
    drop_na(BurnDate) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")))

( lcc_dob_drop_na <- lcc_dob_df %>% drop_na(BurnDate) %>% 
    select(-contains(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .x), ends_with(".x")) %>%
    mutate(fireYr = as.numeric(fireYr)))
```

```{r}
fisl_dob_drop_na
```

```{r}
lc_dob_drop_na
```

```{r}
lcc_dob_drop_na
```


```{r eval=FALSE, include=FALSE}
st_write(fisl_dob_drop_na, "../outputs/dob_modis/fisl_dob_2025-06-10.gpkg")
st_write(lc_dob_drop_na, "../outputs/dob_modis/lc_dob_2025-06-10.gpkg")
st_write(lcc_dob_drop_na, "../outputs/dob_modis/lcc_dob_2025-06-10.gpkg")
```
