---
title: "Organize Climate Data"
author: "Anna Talucci"
date: "2025-08-18"
output: html_document
---



# Overview

Organize Climate data from Climate NA and Climate EU. 

# Packages

```{r}
library(tidyverse)
```

# Data

## Site

```{r}
lc_site = read_csv("../outputs/LCC_cleaned/csv/2025-09-30_LC_Site_fullCombustionID.csv")
fisl_site = read_csv("../outputs/LCC_cleaned/csv/2025-09-30_FISL_Site_fullCombustionID.csv")
```

## CNA

```{r}
fisl_cna = read_csv("../data/combustion_data_layers/FISL/2025-06-02_climateNA_FISL_Normal_1981_2010MSY.csv")

lc_cna = read_csv("../data/combustion_data_layers/LC/2025-06-02_climateNA_LC_Normal_1981_2010MSY.csv")

lcc_ceu = read_csv("../data/2025-08-08_combustion_layer_update/Climate/2025-08-18_climateEU_LCC_Normal_1981_2010MSY.csv")

lcc_cna = read_csv("../data/2025-08-08_combustion_layer_update/Climate/2025-06-02_climateNA_LCC_Normal_1981_2010MSY.csv")
```

# View data

```{r}
lcc_ceu
```

```{r}
sort(names(lcc_ceu))
```

```{r}
names(lcc_cna)
```

# Clean Data

## Function

```{r}
clean_cna_old = function(df){
  df %>% 
  rename(ID = id1, fireYr = id2) %>%
    select(ID,fireYr, lat, long, AHM, bFFP, CMD, CMD05, CMD06, CMD07, CMD08, DD_0, DD_0_05, DD_0_06, DD_0_07, DD_0_08, DD_18, DD18_05, DD18_06, DD18_07, DD18_08, DD18, DD18_sm, DD18_05, DD18_06, DD18_07, DD18_08, DD5, DD5_05, DD5_06, DD5_07, DD5_08, eFFP, EMT, Eref, Eref05, Eref06, Eref07, Eref08, EXT, FFP, MAP, MAR, MAT, MCMT, MSP, MWMT, NFFD, NFFD05, NFFD06, NFFD07, NFFD08, PAS, PAS05, PAS06, PAS07, PAS08, PPT05, PPT06, PPT07, PPT08, Rad05, Rad06, Rad07, Rad08, RH, RH05, RH06, RH07, RH08, SHM, Tave05, Tave06, Tave07, Tave08, TD, Tmax05, Tmax06, Tmax07, Tmax08, Tmin05, Tmin06, Tmin07, Tmin08) %>%
    mutate(CMD_5_8 = ((CMD05 + CMD06 + CMD07 + CMD08)/4),
           DD_0_5_8 = ((DD_0_05 + DD_0_06 + DD_0_07 + DD_0_08)/4),
           DD_18_5_8 = ((DD18_05 + DD18_06 + DD18_07 + DD18_08)/4),
           DD5_5_8 = ((DD5_05 + DD5_06 + DD5_07 + DD5_08)/4),
           Eref_5_8 = ((Eref05 + Eref06 + Eref07 + Eref08)/4),
           NFFD_5_8 = ((NFFD05 + NFFD06 + NFFD07 + NFFD08)/4),
           PAS_5_8 = ((PAS05 + PAS06 + PAS07 + PAS08)/4),
           PPT_5_8 = ((PPT05 + PPT06 + PPT07 + PPT08)/4),
           Rad_5_8 = ((Rad05 + Rad06 + Rad07 + Rad08)/4), 
           RH_5_8 = ((RH05 + RH06 + RH07 + RH08)/4),
           Tave_5_8 = ((Tave05 + Tave06 + Tave07 + Tave08)/4),
           Tmax_5_8 = ((Tmax05 + Tmax06 + Tmax07 + Tmax08)/4),
           Tmin_5_8 = ((Tmin05 + Tmin06 + Tmin07 + Tmin08)/4)) %>%
    select(-contains("05"), -contains("06"), -contains("07"), -contains("08")) %>%
     dplyr::rename_with(~ paste("CNA", .x, sep = "_")) %>%
    rename(ID = CNA_ID)
}
```
select(ID,  AHM, bFFP, CMD, `DD<0`,DD<18"     "DD>18"     "DD>5"     
 [8] "eFFP"      "Elevation" "EMT"       "Eref"      "FFP"       "id1"       "id2"      
[15] "Latitude"  "Longitude" "MAP"       "MAT"       "MCMT"      "MSP"       "MWMT"     
[22] "NFFD"      "PAS"       "PPT_at"    "PPT_sm"    "PPT_sp"    "PPT_wt"    "PPT01"    
[29] "PPT02"     "PPT03"     "PPT04"     "PPT05"     "PPT06"     "PPT07"     "PPT08"    
[36] "PPT09"     "PPT10"     "PPT11"     "PPT12"     "SHM"       "Tave_at"   "Tave_sm"  
[43] "Tave_sp"   "Tave_wt"   "Tave01"    "Tave02"    "Tave03"    "Tave04"    "Tave05"   
[50] "Tave06"    "Tave07"    "Tave08"    "Tave09"    "Tave10"    "Tave11"    "Tave12"   
[57] "TD"        "Tmax_at"   "Tmax_sm"   "Tmax_sp"   "Tmax_wt"   "Tmax01"    "Tmax02"   
[64] "Tmax03"    "Tmax04"    "Tmax05"    "Tmax06"    "Tmax07"    "Tmax08"    "Tmax09"   
[71] "Tmax10"    "Tmax11"    "Tmax12"    "Tmin_at"   "Tmin_sm"   "Tmin_sp"   "Tmin_wt"  
[78] "Tmin01"    "Tmin02"    "Tmin03"    "Tmin04"    "Tmin05"    "Tmin06"    "Tmin07"   
[85] "Tmin08"    "Tmin09"    "Tmin10"    "Tmin11"    "Tmin12"  ) %>%

"> greater than"
"< less than"

CNA_DD_0	degree-days below 0째C (chilling degree days)
CNA_DD5	degree-days above 5째C (growing degree days)
CNA_DD_18	degree-days below 18째C
CNA_DD18	degree-days above 18째C

PPT01, PPT02, PPT03, PPT04, PPT05, PPT06, PPT07, PPT08, PPT09, PPT10, PPT11, PPT12,
Tave01, Tave02, Tave03, Tave04, Tave05, Tave06, Tave07, Tave08, Tave09, Tave10, Tave11, Tave12,
Tmax01, Tmax02, Tmax03, Tmax04, Tmax05, Tmax06, Tmax07, Tmax08, Tmax09, Tmax10, Tmax11, Tmax12,
Tmin01, Tmin02, Tmin03, Tmin04, Tmin05, Tmin06, Tmin07, Tmin08, Tmin09, Tmin10, Tmin11, Tmin12

```{r}
clean_ceu = function(df){
  df %>% 
  rename(ID = id1, DD_0=`DD<0`, DD_18=`DD<18`, DD18=`DD>18`, DD5 = `DD>5`) %>%
    select(ID,  AHM, bFFP, CMD, DD_0, DD_18, DD18, DD5, eFFP, EMT, Eref, FFP, MAP, MAT, MCMT, MSP, MWMT, NFFD, PAS, PPT_at, PPT_sm, PPT_sp, PPT_wt,  SHM, Tave_at, Tave_sm, Tave_sp, Tave_wt,  TD, Tmax_at, Tmax_sm, Tmax_sp, Tmax_wt,  Tmin_at, Tmin_sm, Tmin_sp, Tmin_wt)
}

clean_cna = function(df){
  df %>% 
  rename(ID = id1, fireYr = id2) %>%
    select(ID, fireYr, Latitude, Longitude, AHM, bFFP, CMD, DD_0, DD_18, DD18, DD5, eFFP, EMT, Eref, FFP, MAP, MAT, MCMT, MSP, MWMT, NFFD, PAS, PPT_at, PPT_sm, PPT_sp, PPT_wt,  SHM, Tave_at, Tave_sm, Tave_sp, Tave_wt,  TD, Tmax_at, Tmax_sm, Tmax_sp, Tmax_wt,  Tmin_at, Tmin_sm, Tmin_sp, Tmin_wt)
}
```

## Clean dataframe variables

```{r}
(fisl_cna_clean = clean_cna(fisl_cna) )
( lc_cna_clean = clean_cna(lc_cna) )
( lcc_cna_clean = clean_cna(lcc_cna) )
( lcc_ceu_clean = clean_ceu(lcc_ceu) )
```
# Combine LCC NA and EU to single data frame

```{r}
lcc_ceu_clean
lcc_cna_clean
```
```{r}
unique(lcc_ceu_clean$ID)
```

```{r}
target_id = c("LCC_20", "LCC_21", "LCC_22", "LCC_23", "LCC_24", "LCC_25", "LCC_26", "LCC_27", "LCC_28", "LCC_29", "LCC_55", "LCC_56", "LCC_57", "LCC_58", "LCC_59", "LCC_60", "LCC_61", "LCC_62", "LCC_63", "LCC_64")
```

```{r}
lcc_cna_clean %>% filter(ID %in% target_id)
```
```{r}
( lcc_naonly = lcc_cna_clean %>% filter(AHM != -9999) )
```
```{r}
( lcc_combine = bind_rows(lcc_naonly, lcc_ceu_clean) )
```

# Subset CNA by FISL and LC site level

```{r}
fisl_site
lc_site
```

## FISL

```{r}
fisl_cna_clean %>% filter(ID %in% c('FISL_233', 'FISL_234', 'FISL_235'))
```

```{r}
fisl_site %>% filter(site_ID == "FISL_44")
```
FISL_44	44	Shovel Creek	20	FISL_233, FISL_234, FISL_235
```{r}
( fisl_cna_site = fisl_cna_clean %>%
  filter(fireYr != 2004) %>%
  distinct(Latitude, Longitude, .keep_all = TRUE) %>%
  rename(lat = Latitude, lon = Longitude) %>%
  full_join(fisl_site, by = c("lat", "lon", "fireYr")) %>%
  drop_na() %>%
    select(-fireYr, -lat, -lon, -ID) %>% 
    rename(ID=site_ID) %>% 
    relocate(ID) %>% select(ID:Tmin_wt) )
```

```{r}
fisl_cna_clean %>% filter(ID %in% c('FISL_233', 'FISL_234', 'FISL_235'))
(fisl_44 = fisl_cna_clean %>% 
    filter(ID =='FISL_233') %>%
    mutate(ID = if_else(ID == "FISL_233", "FISL_44", ID)) %>%
    select(-fireYr, -Latitude, -Longitude)  ) 
```

```{r}
( fisl_cna_site1 = bind_rows(fisl_cna_site, fisl_44) )
```


## LC

```{r}
( lc_cna_site = lc_cna_clean  %>%
  mutate(Longitude = case_when(ID %in% c('LC_112', 'LC_113', 'LC_118') ~ -151.458963, TRUE ~ Longitude)) %>%
  distinct(Latitude, Longitude, .keep_all = TRUE) %>%
  rename(lat = Latitude, lon = Longitude) %>%
  full_join(lc_site, by = c("lat", "lon", "fireYr")) %>%
  drop_na() %>%
  select(-fireYr, -lat, -lon, -ID) %>% 
    rename(ID=site_ID) %>% 
    relocate(ID) %>% 
    select(ID:Tmin_wt) 
  )
```


# Save DF as csv

```{r}

write_csv(fisl_cna_site1, "../outputs/2025-08-18_cleaned/csv/2025-10-02_Fisl_site_ClimateNormals.csv")
write_csv(lc_cna_site, "../outputs/2025-08-18_cleaned/csv/2025-10-02_LC_site_ClimateNormals.csv")

write_csv(lcc_combine, "../outputs/2025-08-18_cleaned/csv/2025-08-20_LCC_ClimateNormals.csv")
```